DRAFT \\ From physical principles to classical and quantum \\ Hamiltonian and Lagrangian particle mechanics

### 2016-05-18 2015CompleteDerivation/completeDerivation.tex
Minor improvement after Monday discussion

\begin{abstract}
\textbf{This is manuscript is a work in progress.} Ideas are constantly reshaped to find more precise and elegant arguments. It is provided as is to stimulate discussion.

@@

Given that scientific reductionism (i.e. the idea of reducing physical systems and interactions to the sum of their constituent parts in order to make them easier to study) is at the heart of fundamental physics, we now explore how to characterize a system in terms of its components. That is, we want to study the relationship between the state space $\mathcal{C}$ of a composite system and the state space $\mathcal{S}$ of its parts. In general, this is quite a complicated thing to do, which requires intimate knowledge of the system at hand. So we simplify our problem and study a system made of infinitesimal homogeneous parts. What we'll find is that under the additional assumption that the system is infinitesimally reducible (i.e. its state is equivalent to the state of its parts) and that each part undergoes deterministic and reversible evolution, the motion is suitably described by the standard framework of classical Hamiltonian mechanics.

@@

\begin{prop}\label{reducible_state_space}
The state space $\mathcal{C}$ for decomposable systems made of a homogeneous material is an additive abelian (i.e. commutative) group.
\end{prop}

@@

We claim $\mathcal{C}$ is an additive group. As we want to describe changes during the evolution (i.e. $\hat{\mathcal{c}} = \mathcal{c} +\delta \mathcal{c}$) we introduce an inverse $- : \mathcal{C} \rightarrow \mathcal{C}$ such that $\mathcal{c} + ( - \mathcal{c}) = 0 \forall \mathcal{c} \in \mathcal{C}$. Such inverse may introduce objects that do not properly represent a physical state, but a state change. A change of a physically distinguishable object is itself physically distinguishable therefore $\mathcal{C}$ is still a Hausdorff and second countable topological manifold as per \ref{topologically_distinguishable}. We'll still call $\mathcal{C}$ state space committing an abuse of terminology.
\end{justification}

@@

\begin{prop}\label{classical_vector space}
The state space $\mathcal{C}$ for a homogeneous classical (i.e. infinitesimally reducible) material is a vector space over $\mathbbm{R}$ isomorphic to a subspace of $C(\mathcal{S}) = \{\rho:\mathcal{S} \rightarrow \mathbbm{R} \; | \; \rho$ is continuous$\}$, where $\mathcal{S}$ is the state space of an infinitesimal amount of material.
\end{prop}

@@

\begin{prop}\label{integration}
	Let $\mathcal{S}$ be the state space for the particles of a homogeneous classical material. Let each state $\mathcal{s} \in \mathcal{S}$ be fully identified by a finite number of independent state variables. Then $\mathcal{S}$ is endowed with a Borel measure $\mu$. The state space $\mathcal{C}$ is isomorphic to the space of continuous functions that are integrable over any subset of $\mathcal{S}$. That is: $\mathcal{C} \cong C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ where $L^1(\mathcal{S}, \mu) = \{ \rho : \mathcal{S} \rightarrow \mathbbm{R} \; | \; \int_{\mathcal{S}} |\rho| d\mu < \infty \}$.
\end{prop}

@@

@@

We recognize the familiar set of Hamilton's equations. They are the set of equation to describe the deterministic and reversible motion of the infinitesimal parts of an infinitesimally reducible homogeneous material. In other words: \textbf{the forces that conserve energy (i.e. the value of a sutiable Hamiltonian) are exactly the ones that provide deterministic and reversible motion}. The challenge in their derivation mainly lies in the necessary use of different branches of mathematics (e.g. topology, measure theory, differential geometry, symplectic geometry). The conceptual meaning, on the other hand, is hopefully straight forward.

>>>>

\begin{abstract}
\textbf{This manuscript is a work in progress.} Ideas are constantly reshaped to find more precise and elegant arguments. It is provided as is to stimulate discussion.

@@

Given that scientific reductionism (i.e. the idea of reducing physical systems and interactions to the sum of their constituent parts in order to make them easier to study) is at the heart of fundamental physics, we now explore how to characterize a system in terms of its components. That is, we want to study the relationship between the state space of a composite system and the state space of its parts. In general, this is quite a complicated thing to do, which requires intimate knowledge of the system at hand. So we simplify our problem and study a system made of infinitesimal homogeneous parts. What we'll find is that under the additional assumption that the system is infinitesimally reducible (i.e. its state is equivalent to the states of its parts) and that each part undergoes deterministic and reversible evolution, the motion is suitably described by the standard framework of classical Hamiltonian mechanics.

@@

\begin{prop}\label{reducible_state_space}
The state space $\mathcal{C}$ for a decomposable homogeneous material is an additive abelian (i.e. commutative) group.
\end{prop}

@@

We claim $\mathcal{C}$ is an additive group. As we want to describe changes during the evolution (i.e. $\hat{\mathcal{c}} = \mathcal{c} +\delta \mathcal{c}$) we introduce an inverse $- : \mathcal{C} \rightarrow \mathcal{C}$ such that $\mathcal{c} + ( - \mathcal{c}) = 0 \forall \mathcal{c} \in \mathcal{C}$. Such inverse includes objects that do not properly represent a physical state, but a state change. A change of a physically distinguishable object is itself physically distinguishable therefore $\mathcal{C}$ is still a Hausdorff and second countable topological manifold as per \ref{topologically_distinguishable}.
\end{justification}

@@

\begin{prop}\label{classical_vector space}
The state space $\mathcal{C}$ for a homogeneous classical (i.e. infinitesimally reducible) material is a vector space over $\mathbbm{R}$ isomorphic to a subspace of $C(\mathcal{S}) \equiv \{\rho:\mathcal{S} \rightarrow \mathbbm{R} \; | \; \rho$ is continuous$\}$, where $\mathcal{S}$ is the state space of an infinitesimal amount of material.
\end{prop}

@@

\begin{prop}\label{integration}
	Let $\mathcal{S}$ be the state space for the particles of a homogeneous classical material. Let each state $\mathcal{s} \in \mathcal{S}$ be fully identified by a finite number of independent state variables. Then $\mathcal{S}$ is endowed with a Borel measure $\mu$. The state space $\mathcal{C}$ is isomorphic to the space of Lebesgue integrable continuous functions over $\mathcal{S}$. That is: $\mathcal{C} \cong C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ where $L^1(\mathcal{S}, \mu) = \{ \rho : \mathcal{S} \rightarrow \mathbbm{R} \; | \; \int_{\mathcal{S}} |\rho| d\mu < \infty \}$.
\end{prop}

@@

    Being able to track parts at a given scale is not even enough. The parts are infinitesimal, the limit of a process of infinite recursive subdivision, so we must be able to track at any scale. Therefore the assumption can only be validated experimentally to the smallest precision available.

@@

We recognize the familiar set of Hamilton's equations. They are the set of equation to describe the deterministic and reversible motion of the infinitesimal parts of an infinitesimally reducible homogeneous material. In other words: the forces that conserve energy (i.e. the value of a suitable Hamiltonian) are exactly the ones that provide deterministic and reversible motion. The challenge in their derivation mainly lies in the necessary use of different branches of mathematics (e.g. topology, measure theory, differential geometry, symplectic geometry). The conceptual meaning, on the other hand, is hopefully straight forward.
